---
title: "Fine-mapping analysis pipeline by `finemapr`"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", # NA
  dev = "png", dev.args = list(type = "cairo"), dpi = 92, fig.path = "figures/", 
  fig.width = 4, fig.height = 4,
  results = 'markup', tidy = F, message = T, warning = T, echo = T
)
```

# About `finemapr`

`finemapr` is an R package that provides an interface to fine-mapping tools:

- FINEMAP http://www.christianbenner.com/
- CAVIAR https://github.com/fhormoz/caviar
- PAINTOR https://github.com/gkichaev/PAINTOR_V3.0

By using `finemapr`, your input files are automatically prepared for each tool, the analysis workflow is tool-independent; and exploration of fine-mapping results is powered by R in printing/plotting/data export.


## Tool-independent scheme of analysis workflow

```r
# set up
options(finemapr_<tool> = "<path to fine-mapping tool>")

# read input files
my_zscores <- read_zscores("<my_scores.tab>")
my_ld <- read_ld("<my_ld.tab>")

# run analysis
out <- run_<tool>(my_zscores, my_ld, args = "<custom arguments>")

# explore results
print(out)
head(out$snp) # main table of results
plot(out)

# export results
write.table(out$snp, "<my_results.tab>")
```

# Installation

The user needs to download and install a fine-mapping tool before the analysis.
An example of installation commands used in `finemapr` by default is given [here](https://github.com/variani/finemapr/blob/master/misc/install-finemaping-tools.md). 

After installing, for example, the FINEMAP tool, the user specify for `finemapr` where the tool is located:

```{r, eval = F}
options(finemapr_finemap = "~/apps/finemap/finemap")
```

# Load packages

We load packages for the analysis conducted in this document.

```{r inc}
library(devtools)
load_all("~/git/variani/finemapr")

library(magrittr)
library(dplyr)  
library(ggplot2)  
theme_set(theme_linedraw())
```

# Example data

We load example data copied from the FINEMAP website (http://www.christianbenner.com/).
This simulated dataset has two causal variants `rs15` and `rs47`.

```{r example_data}
file1_z <- system.file("extdata/region1.z", package = "finemapr")
file1_ld <- system.file("extdata/region1.ld", package = "finemapr")

z1 <- read_zscore(file1_z)
ld1 <- read_ld(file1_ld, snps = z1$snp)
n1 <- 5363
```

## Explore z-scores

Top 5 z-scores:

```{r top5, results = "asis"}
z1 %>% arrange(-abs(zscore)) %>% head(5) %>% kable(digits = 1)
```

```{r plot_zscore}
ggplot(z1, aes(zscore)) + geom_histogram()
```

```{r plot_pval}
mutate(z1, pval = pchisq(zscore^2, df = 1, lower.tail = FALSE)) %>%
  ggplot(aes(pval)) + geom_histogram()
```

# Run tools

## Run FINEMAP

```{r run_finemap}
options(finemapr_finemap = "~/apps/finemap/finemap")

out_finemap <- run_finemap(z1, ld1, n1, args = "--n-causal-max 3")
```

```{r print_finemap}
print(out_finemap)
```

```{r plot_finemap, fig.width = 6, fig.height = 9}
plot(out_finemap, label_size = 3, grid_ncol = 1)
```

## Run CAVIAR

```{r run_caviar}
options(finemapr_caviar = "~/apps/caviar/CAVIAR")

out_caviar <- run_caviar(z1, ld1, args = "-c 3")
```

```{r print_caviar}
print(out_caviar)
```

```{r plot_caviar, fig.width = 6, fig.height = 3}
plot(out_caviar, label_size = 3)
```

## Run PAINTOR

```{r run_paintor}
options(finemapr_paintor = "~/apps/paintor/PAINTOR")

out_paintor <- run_paintor(z1, ld1, n1, args = "-enumerate 3")
```

```{r print_paintor}
print(out_paintor)
```

```{r plot_paintor, fig.width = 6, fig.height = 3}
plot(out_paintor, label_size = 3)
```

# Conclusions

All three fine-mapping tools estimated the poterior causal probabilities of the two variants, `rs15` and `rs47`, very close to 1.
